---
title: "molkart.misty_neighbourhood_analysis"
author: "FloWuenne"
date: "2023-08-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(Seurat)
#options(Seurat.object.assay.version = 'v5')
library(tidyverse)
library(scales)
library(pals)
library(patchwork)
library(vitessceR)
library(Nebulosa)
library(mistyR)
library(ClusterR)
library(future)

plan("multisession", workers=4)

cellular_neighborhoods <- function(sample.cells, sample.pos, k,threshold){
  misty.views  <- create_initial_view(sample.cells) %>% add_juxtaview(sample.pos, neighbor.thr = threshold)
  clust <- KMeans_rcpp(misty.views[[3]], k)
  return(clust$clusters) 
}

source("./code/functions.R")
```


# Introduction

In this analysis, we will estimate cell-type abundance changes across the dataset, use misty to investigate the changing tissue architecture and microenvironments and check neighbourhood compositions.

# Load data

```{r}
## If the object has already been computed
seurat_object <- readRDS(file = "./output/mol_cart/molkart.harmony_seurat_object.rds")
```


# Misty results

## Niche analysis using Misty

```{r}
library(rlist)
all.data <- read_tsv("./output/mol_cart/molcart.misty_celltype_table.tsv")

samples <- all.data %>%
  pull(sample_ID) %>%
  unique()
cts <- all.data %>%
  pull(anno_cell_type_lv2) %>%
  unique()

cts.names <- make.names(cts, allow_ = FALSE)

sample <- "sample_2d_r2_s1"
sample.cells <- all.data %>%
    filter(sample_ID == sample) %>%
    pull(anno_cell_type_lv2) %>%
    map(~ .x == cts) %>%
    list.rbind() %>%
    `colnames<-`(cts.names) %>%
    as_tibble()

sample.pos <- all.data %>%
    filter(sample_ID == sample) %>%
    select(X_centroid, Y_centroid)

niche_clusters <- cellular_neighborhoods(sample.cells,sample.pos, k = 5, thresh = NULL)
```


```{r}
sample_seurat <- subset(seurat_object,sample_ID == sample)
cell_types <- sample_seurat@meta.data$anno_cell_type_lv2
coord_exp <- cbind(sample.pos,niche_clusters,cell_types)

all_niches <- ggplot(coord_exp,aes(Y_centroid,X_centroid, color = as.factor(niche_clusters))) +
  geom_point() +
  scale_color_brewer(palette = "Set3")
all_niches

niche_plot <- ggplot(coord_exp,aes(Y_centroid,X_centroid)) +
  geom_point(data = subset(coord_exp,niche_clusters != 2),color ="darkgrey") +
  geom_point(data = subset(coord_exp,niche_clusters == 2),color = "purple")
niche_plot

niches_facet <- ggplot(coord_exp,aes(Y_centroid,X_centroid)) +
  geom_point(aes(color = as.factor(niche_clusters))) +
  facet_wrap(~ niche_clusters) +
  theme(legend.position = "none") + 
  scale_color_brewer(palette = "Set3")

ct_per_niche <- coord_exp %>%
  group_by(niche_clusters,cell_types) %>%
  tally()

ct_niches <- ggplot(ct_per_niche,aes(niche_clusters,n,fill = cell_types)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set3")
ct_niches

niches_facet + ct_niches
```


# Seurat niche analysis

## Density expression plot

```{r}
seurat_object_d2_r1 <- subset(seurat_object,sample_ID == "sample_2d_r1_s1")
seurat_object_d2_r2 <- subset(seurat_object,sample_ID == "sample_2d_r2_s1")

coordinates <- seurat_object_d2_r2@meta.data %>%
  select(X_centroid,Y_centroid)
seurat_object_d2_r2@reductions$pca@cell.embeddings[,1] <- coordinates$Y_centroid
seurat_object_d2_r2@reductions$pca@cell.embeddings[,2] <- coordinates$X_centroid

plot_density(seurat_object_d2_r2, reduction = "pca", feature = "Ccr2")
```

```{r}
exp_matrix <- seurat_object_d2_r2@assays$SCT@data[c("Nppb"),]
coord_exp <- cbind(coordinates,exp_matrix)

ggplot(coord_exp,aes(Y_centroid,X_centroid)) +
  geom_point(data = subset(coord_exp, exp_matrix == 0), color ="darkgrey") +
  geom_point(data = subset(coord_exp, exp_matrix > 0), aes(color = exp_matrix)) +
  scale_color_viridis_b()
```

## Plot Nppa, Nppb and CM layers

```{r}
seurat_object_d2_r1 <- subset(seurat_object,sample_ID == "sample_2d_r1_s1")
seurat_object_d2_r2 <- subset(seurat_object,sample_ID == "sample_2d_r2_s1")
meta <- seurat_object_d2_r2@meta.data

ggplot(meta,aes(Y_centroid,X_centroid)) +
  geom_point(data = subset(meta, !anno_cell_type_lv2 %in% c("Cardiomyocytes_Nppa+","Cardiomyocytes_Nppb+","Cardiomyocytes","Myeloid_cells")),
             color = "darkgrey") +
  geom_point(data = subset(meta, anno_cell_type_lv2 %in% c("Cardiomyocytes_Nppa+")),
             color = "magenta") +
  geom_point(data = subset(meta, anno_cell_type_lv2 %in% c("Cardiomyocytes_Nppb+")),
             color = "orange") +
  geom_point(data = subset(meta, anno_cell_type_lv2 %in% c("Cardiomyocytes")),
             color = "darkgreen") +
    geom_point(data = subset(meta, anno_cell_type_lv2 %in% c("Myeloid_cells")),
             color = "cyan")

```


# Add spatial coordinates as spatial data assay to access spatial functions

```{r}
cell_coordinates <- cell_coordinates[, .(cell_id = V1, x = Y, y = X,
  x_tile = Y, y_tile = X, size = Width * Height)]
```


```{r}
seurat_object_d2_r1 <- subset(seurat_object,sample_ID == "sample_2d_r1_s1")
sample_coordinates <- 

seurat_object_d2_r1@images <- list()
seurat_object_d2_r1@images["cen"] <- CreateFOV(
  coords = sample_coordinates,
  type = "centroids",
  nsides = 0L,
  radius = 1L,
  theta = 0L,
  molecules = transcripts, # Only for visualising the raw transcripts, can be skipped
  assay = "Spatial",
  key = NULL,
  name = NULL)
```



# General tissue plots



## Plot Specific distributions of cells on tissue

## Level 2

```{r}
expression_plot_list <- list()

samples <- c("sample_control_r1_s1","sample_4h_r1_s1",
             "sample_2d_r1_s1","sample_4d_r1_s1",
             "sample_control_r2_s1","sample_4h_r2_s2",
             "sample_2d_r2_s1","sample_4d_r2_s1")

cluster_of_int <- "Endocardial_cell"
for(this_sample in samples){
    pt_size <- 0.6
    sample_object <- subset(seurat_object,sample_ID == this_sample)
    meta <- sample_object@meta.data
    
    highlight_plot <- ggplot(meta,aes(Y_centroid,X_centroid)) +
      geom_point(data = subset(meta,!seurat_clusters %in% cluster_of_int),color = "darkgrey", size = pt_size) +
      # geom_point(data = subset(meta,seurat_clusters %in% cluster_of_int),aes(color = seurat_clusters), size = pt_size) +
      geom_point(data = subset(meta,anno_cell_type_lv2 == cluster_of_int),color = "purple", size = 1) +
      theme_classic() +
      labs(x = "Spatial 1",
           y = "Spatial 2") +
      theme(axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "right")
    
    expression_plot_list[[this_sample]] <- highlight_plot
}
  
time_plot <- wrap_plots(expression_plot_list, nrow = 2, ncol = 4)  + plot_layout(guides = 'collect') + plot_annotation(cluster_of_int,theme=theme(plot.title=element_text(hjust=0.5)))
  print(time_plot)
  
#filename <- paste("./output/molkart_cell_types/",cluster_of_int,".cell_type_distributions.png",sep = "")
#save_plot(time_plot,
#         filename = filename,
#         base_height = 6)
```

## Level 3

```{r}
expression_plot_list <- list()

samples <- c("sample_control_r1_s1","sample_4h_r1_s1",
             "sample_2d_r1_s1","sample_4d_r1_s1",
             "sample_control_r2_s1","sample_4h_r2_s2",
             "sample_2d_r2_s1","sample_4d_r2_s1")

cluster_of_int <- "yes"
for(this_sample in samples){
    pt_size <- 0.6
    sample_object <- subset(seurat_object,sample_ID == this_sample)
    meta <- sample_object@meta.data
    
    highlight_plot <- ggplot(meta,aes(Y_centroid,X_centroid)) +
      geom_point(data = subset(meta,!seurat_clusters %in% cluster_of_int),color = "darkgrey", size = pt_size) +
      # geom_point(data = subset(meta,seurat_clusters %in% cluster_of_int),aes(color = seurat_clusters), size = pt_size) +
      geom_point(data = subset(meta,ccr2_pos == cluster_of_int),color = "purple", size = 1) +
      theme_classic() +
      labs(x = "Spatial 1",
           y = "Spatial 2") +
      theme(axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "right")
    
    expression_plot_list[[this_sample]] <- highlight_plot
}
  
time_plot <- wrap_plots(expression_plot_list, nrow = 2, ncol = 4)  + plot_layout(guides = 'collect') + plot_annotation(cluster_of_int,theme=theme(plot.title=element_text(hjust=0.5)))
  print(time_plot)
  
#filename <- paste("./output/molkart_cell_types/",cluster_of_int,".cell_type_distributions.png",sep = "")
#save_plot(time_plot,
#         filename = filename,
#         base_height = 6)
```

