---
title: "Process quantifications from Mesmer masks"
author: "Florian Wuennemann"
date: "2024-01-28"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(SeuratData)
library(SeuratDisk)
library(data.table)
library(harmony)
library(here)
library(Nebulosa)
library(tidyverse)
library(GGally)
library(Matrix)
library(sf)
library(sfheaders)
library(viridis)

source("../../code/functions.R")
```

In this markdown, we wil use the cell quantifications from the Mesmer whole cell model. This model seems to perform better in detecting nuclei, but misses a lot of larger cells without nuclear stain.

### Merging seurat objects

```{r}
final_samples <- c("sample_control_r1_s1","sample_control_r2_s1",
                   "sample_4h_r1_s1","sample_4h_r2_s2",
                   "sample_2d_r1_s1","sample_2d_r2_s1",
                   "sample_4d_r1_s1","sample_4d_r2_s1")

seurat_objects <- list()
norm_list <- list()

## SCTransform based clustering across all samples
sample_dir <- "../../../results/nf-core_molkart/spot2cell"
getwd()
seurat_list <- list()
segmethod_samples <- list.files(sample_dir)
for(sample in segmethod_samples){
    sample_ID <- strsplit(sample,"\\.")[[1]][1]
    sample_ID <- strsplit(sample_ID,split="_")[[1]][-1]
    sample_ID <- paste(sample_ID[1:length(sample_ID) -1],collapse = "_")
    if(endsWith(sample,"_mesmer.csv") & sample_ID %in% final_samples){
      print(sample_ID)
      sample_quant <- fread(paste(sample_dir,sample,sep="/"))
      seurat_object <- create_seurat_sctransform_mcquant(sample_quant,sample_ID)
      # norm_list[[sample_ID]] <- seurat_object@assays$RNA@layers$counts / seurat_object@meta.data$Area
      seurat_list[[sample_ID]] <- seurat_object
  }
}

resolve_object <- merge(seurat_list[[1]], y = c(seurat_list[-1]), 
                        project = "molkart_MI", merge.data = TRUE)
resolve_object <- JoinLayers(resolve_object)
```



# Process data

## Plot distribution of transcript counts and mask properties

```{r}
## Set order of timepoints
resolve_object@meta.data$timepoint <- factor(resolve_object@meta.data$timepoint,
                               levels = c("control","4h","2d","4d"))


## Show distribution of QC measures before filtering outliers
VlnPlot(resolve_object, features = c("nCount_RNA","nFeature_RNA","Area","Eccentricity","Solidity"),
        group.by = "timepoint", pt.size = 0, cols = time_palette)

prefiltered_cells <- nrow(resolve_object@meta.data)
print(paste(prefiltered_cells,"cells before filtering"))
```

## Filter outlier cells

```{r}
## Before filtering cells, let's see if there are any specific outliers
meta_props <- resolve_object@meta.data %>%
  select(Area,MajorAxisLength,MinorAxisLength,Eccentricity,Solidity,Extent,nCount_RNA,timepoint)

ggpairs(meta_props, 
        mapping = aes(color = timepoint),
        columns = c("Area", "Solidity", "Extent","nCount_RNA"))
```


```{r}
resolve_object <- subset(resolve_object, 
                         subset = 
                           nCount_RNA <= 4000 & 
                           nCount_RNA >= 20 &
                           Extent > 0.25 &
                           Solidity > 0.75)

## How many cells are left after filtering?
post_filter_cells <- nrow(resolve_object@meta.data)
```

```{r}
## Let's check the same correlations after filtering
meta_props <- resolve_object@meta.data %>%
  select(Area,MajorAxisLength,MinorAxisLength,Eccentricity,Solidity,Extent,nCount_RNA,timepoint)

ggpairs(meta_props, 
        mapping = aes(color = timepoint),
        columns = c("Area", "Solidity", "Extent","nCount_RNA"))
```

```{r}
## Show distribution of QC measures after filtering outliers
VlnPlot(resolve_object, features = c("nCount_RNA","nFeature_RNA","Area","Eccentricity","Solidity"),
        group.by = "timepoint", pt.size = 0, cols = time_palette)
```


## Integrate samples using SCTransform and Harmony

```{r}
## Let's generate a batch variable
#resolve_object$batch <- paste(resolve_object$timepoint,resolve_object$slide,sep="_")
resolve_object[["RNA"]] <- split(resolve_object[["RNA"]], f = resolve_object$timepoint)
```

```{r}
## Normalize and scale data using sctransform v2
resolve_object <- SCTransform(resolve_object, 
                              assay = "RNA")
```

```{r}
# Perform PCA
npcs_to_use <- 30
resolve_object <- RunPCA(resolve_object, npcs = npcs_to_use, verbose = FALSE, approx=FALSE)
ElbowPlot(resolve_object, ndims = npcs_to_use)
```


```{r}
harmony_object <- IntegrateLayers(
  object = resolve_object, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "harmony",
  assay = "SCT", verbose = FALSE
)

harmony_object <- FindNeighbors(harmony_object, reduction = "harmony", dims = 1:npcs_to_use)
harmony_object <- FindClusters(harmony_object, resolution = 0.4, cluster.name = "harmony_clusters")
harmony_object <- RunUMAP(harmony_object, reduction = "harmony", dims = 1:npcs_to_use, reduction.name = "umap")
p1 <- plot_density(harmony_object, features= c("Npr3"), size = 0.5)
p2 <- DimPlot(harmony_object, reduction = "umap", label = TRUE, repel = TRUE,raster=FALSE)
p1+p2
```


## Plot Harmony integration results

```{r}
## Plot UMAP with cells labeled by sampled time point and by cluster assigned using nearest-neighbor analysis
p1 <- DimPlot(harmony_object, reduction = "umap", group.by = "timepoint",raster=FALSE, 
              cols = time_palette, alpha = 0.5)
p2 <- DimPlot(harmony_object, reduction = "umap", label = TRUE, repel = TRUE,raster=FALSE)
p1+p2
```


```{r}
## Plot UMAP next to a barplot
meta_time <- harmony_object@meta.data %>%
  group_by(seurat_clusters, timepoint) %>%
  tally() %>%
  mutate("frac" = n / sum(n))

time_bar <- ggplot(meta_time,aes(seurat_clusters,frac, fill = timepoint)) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(values = time_palette)

combined_time_plot <- p2 + time_bar & 
  theme(plot.background = element_rect(fill = 'white'),
        panel.background = element_rect(fill = 'white'))

# save_plot(combined_time_plot,
#           file = "./plots/molkart.umap_time.png",
#           base_height = 10)
```

```{r}
## Plot Nebulosa plots for QC measures to highlight cell clusters with high number of transcripts
plot_density(harmony_object, features= c("Area","nCount_RNA","Solidity","Extent"))
```

```{r}
## Plot the number of cells per cluster and color by cluster ID
cell_numbers <- harmony_object@meta.data %>%
  group_by(seurat_clusters) %>%
  tally()

ggplot(cell_numbers,aes(seurat_clusters,n)) +
  geom_bar(stat = "identity",aes(fill = seurat_clusters)) +
  theme(legend.position = "None")
```

## Identify marker genes for cell clusters

```{r}
harmony_object <- PrepSCTFindMarkers(harmony_object, assay = "SCT", verbose = TRUE)
harmony_markers <- FindAllMarkers(harmony_object, logfc.threshold = 0.5, only.pos = TRUE)
```

```{r}
## Write marker table to file
# write.table(harmony_markers,
#             file = "./output/molkart/molkart.cell_type_markers.csv",
#             sep = ",",
#             row.names = TRUE,
#             col.names = TRUE,
#             quote = FALSE)
```


```{r}
p1 <- plot_density(harmony_object, features= c("Npr3"), size = 0.5)
p2 <- DimPlot(harmony_object, reduction = "umap", label = TRUE, repel = TRUE,raster=FALSE)
p1+p2
```


## Transfer labels from snRNA-seq (Calcagno et al. 2020)

```{r}
calcagno_et_al <- LoadH5Seurat("../public_data/Calcagno_et_al_NatCardioVasc_2022/reprocessed_data/Calcagno2022_int_logNorm_annot.h5Seurat")
```

```{r}
calcagno_et_al <- SCTransform(calcagno_et_al, assay = "RNA")
```

```{r}
anchors <- FindTransferAnchors(reference = calcagno_et_al, query = harmony_object, normalization.method = "SCT", npcs = 50, recompute.residuals = FALSE)
predictions.assay_lvl1 <- TransferData(anchorset = anchors, refdata = calcagno_et_al$level_1, prediction.assay = FALSE, weight.reduction = "pcaproject", dims = NULL)
predictions.assay_lvl2 <- TransferData(anchorset = anchors, refdata = calcagno_et_al$level_2, prediction.assay = FALSE, weight.reduction = "pcaproject", dims = NULL)
predictions.assay_lvl3 <- TransferData(anchorset = anchors, refdata = calcagno_et_al$level_3, prediction.assay = FALSE, weight.reduction = "pcaproject", dims = NULL)
predictions.orig_label <- TransferData(anchorset = anchors, refdata = calcagno_et_al$orig_label, prediction.assay = FALSE, weight.reduction = "pcaproject", dims = NULL)
```


```{r}
## Add labels to Harmony object and plot Umap
harmony_object@meta.data$predicted_ct_lvl1 <- predictions.assay_lvl1$predicted.id
harmony_object@meta.data$predicted_ct_lvl2 <- predictions.assay_lvl2$predicted.id
harmony_object@meta.data$predicted_ct_lvl3 <- predictions.assay_lvl3$predicted.id
harmony_object@meta.data$orig_label_calcagno <- predictions.orig_label$predicted.id
```


```{r}
calcagno_lvl1 <- DimPlot(harmony_object, reduction = "umap", group.by = "predicted_ct_lvl1",raster=FALSE, label = TRUE)
calcagno_lvl2 <- DimPlot(harmony_object, reduction = "umap", group.by = "predicted_ct_lvl2",raster=FALSE, label = TRUE)
calcagno_lvl3 <- DimPlot(harmony_object, reduction = "umap", group.by = "predicted_ct_lvl3",raster=FALSE, label = TRUE)
DimPlot(harmony_object, reduction = "umap", group.by = "orig_label_calcagno",raster=FALSE, label = TRUE)

calcagno_lvl1 | calcagno_lvl2 | calcagno_lvl3
```


## Use manual tissue annotation to improve endocardial cell labeling

Endocardial cells are very small and thin and form a monolayer at the inner lining of the heart. We noticed that our marker Npr3 does not always reliably label endocardial cells and therefore will use manual annotation of the endocardial tissue regions to improve annotation of endocardial cells.


```{r}
## Function to use GeoJson mask from QuPath with Seurat metadata to check cell overlap to manual annotation region
check_mask_overlap <- function(polygon_sf,metadata,img_height,anno_name){
  points_sf <- st_as_sf(metadata, coords = c("X_centroid", "Y_centroid"), crs = NA)
  polygon_sf <- st_set_crs(polygon_sf, NA)
  
  mirrored_geometries <- st_geometry(polygon_sf) %>% 
    st_coordinates() %>%
    as.data.frame() %>%
    mutate("Y" = img_height - Y) %>%
    select(X,Y,L2)
  mirrored_geometries <- as.data.frame(mirrored_geometries)
  
  polygon_list <- split(mirrored_geometries[,c("X", "Y")], mirrored_geometries$L2)
  closed_polys <- list()
  # Iterate over polygons in polygon_list and add the first row to the end of each polygon
  for (i in 1:length(polygon_list)) {
    closed_polys[[i]] <- as.matrix(rbind(polygon_list[[i]], polygon_list[[i]][1,]))
  }
  
  # Convert the list of polygons to sf objects
  sf_objects <- lapply(closed_polys, function(polygon) {
    st_polygon(list(polygon))
  })
  
  # Create a MULTIPOLYGON object
  multipolygon <- st_multipolygon(sf_objects)
  
  ## Check which points fall withing the annotation
  points_within_polygons <- st_within(points_sf, multipolygon, sparse = FALSE)
  metadata[anno_name] <- points_within_polygons
  return(metadata)
}
```

```{r}
## Check which cells fall in the spatial endocardial region for each sample
all_vals <- list()
for(sample_id in unique(harmony_object@meta.data$sample_ID)){
  print(sample_id)
  image_sizes <- fread("./references/molkart.image_dimensions.csv")
  img_height <- subset(image_sizes,sample_ID == sample_id)$height
  
  polygon_sf <- st_read(paste("/Users/florian_wuennemann/1_Projects/MI_project/results/QuPath_annotations/molkart.heart_regions/geojson/",sample_id,".stack.geojson",sep=""))
  sample_meta <- subset(harmony_object@meta.data,sample_ID == sample_id)
  sample_vals <- list(polygon_sf,sample_meta,img_height)
  all_vals <- append(all_vals,list(sample_vals))
  #
}
#test <- check_mask_overlap(polygon_sf,sample_meta,img_height)
meta_list <- Map(function(x) check_mask_overlap(x[[1]], x[[2]], x[[3]],"in_endo"), all_vals)

# Combine all data frames into one
new_full_meta <- do.call(rbind, meta_list)
```

```{r}
## Which seurat cluster corresponds to endocardial cells (Npr3 positive)?
endocardial_cluster <- 20

## How many cells are in the endocardial cluster?
n_endo <- nrow(subset(new_full_meta,seurat_clusters == endocardial_cluster))

## How many of the transcriptionally defined endocardial cell cluster are spatially within the endocardium?
n_endo_spatial <- nrow(subset(new_full_meta,seurat_clusters == endocardial_cluster & in_endo == TRUE))

## How many segmented cells are total in the annotated regions?
anno_region_cells <- nrow(subset(new_full_meta, in_endo == TRUE))

## Let's classify cells that are Pecam1 positive within the endocardial region as endocardial (even if they are not Npr3 positive)
## add Pecam1 expression 
pecam1_exp <- harmony_object@assays$SCT@data["Pecam1",]
new_full_meta$Pecam1 <- pecam1_exp
npr3_exp <- harmony_object@assays$SCT@data["Npr3",]
new_full_meta$Npr3 <- npr3_exp

## How many cells are Pecam1 positive within the endocardial region?
pecam1_pos <- subset(new_full_meta,Pecam1 > 0 & in_endo == TRUE)

## Add another annotation column to metadata to classify all cells with Pecam1 > 0 within the endocardial region as endocardial cells
new_full_meta <- new_full_meta %>%
  mutate("spatial_endoc" = if_else((in_endo == TRUE & Pecam1 > 0) | (in_endo == TRUE & seurat_clusters == endocardial_cluster), TRUE,FALSE))

harmony_object@meta.data <- new_full_meta
```

### Plot results of spatial annotation for endocardial cells 

```{r}
## Plot cells that are transcriptionally labeled as endocardial cells
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey") +
  geom_point(data = subset(new_full_meta,seurat_clusters == endocardial_cluster),color = "red") +
  facet_wrap(~ sample_ID)

## Plot 
## 1: blue - originally transcriptionally labeled endocardial cells but not in endocardial region. 
## 2: red - transcriptionally labeled endocardial cells that were in endocardial region
## 3: purple - Cells that were not clustered with Npr3 positive cells but spatially fall within endocardium and are Pecam1 or Npr3 positive.
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey", size = 0.5) +
  geom_point(data = subset(new_full_meta,spatial_endoc == TRUE & seurat_clusters == endocardial_cluster),color = "red", size = 0.75) +
  geom_point(data = subset(new_full_meta,spatial_endoc == TRUE & seurat_clusters != endocardial_cluster),color = "purple", size = 0.75) +
    geom_point(data = subset(new_full_meta,spatial_endoc == FALSE & seurat_clusters == endocardial_cluster),color = "blue", size = 0.75) +
  facet_wrap(~ sample_ID)

## Plot all cells that were considered "in endocardial region"
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey") +
  geom_point(data = subset(new_full_meta,spatial_endoc == TRUE )) +
  facet_wrap(~ sample_ID)
```

### Level 2 annotation - Sub-Cell & cell states

```{r}
## Annotate cell clusters with more detaile
## Lvl2

seurat_clusters <- harmony_object@active.ident

## Cluster level 2
new.cluster.ids <- c("Cardiomyocytes",  #0
                     "Endothelial_cells",  #1
                     "Cardiac_fibroblasts", #2
                     "Endothelial_cells", #3
                     "Myeloid_cells", #4
                     "Cardiomyocytes_Nppa+", #5 
                     "Cardiac_fibroblasts", #6
                     "Cardiomyocytes", #7
                     "Cardiac_fibroblasts", #8
                     "Cardiomyocytes", #9
                     "EndMA+", #10 
                     "Cardiomyocytes" , #11
                     "Myeloid_cells", #12
                     "Pericytes", #13
                     "Cardiomyocytes_Nppb+", #14
                     "Cardiac_fibroblasts_Il6+", #15
                     "Cardiomyocytes", #16
                     "Smooth_muscle_cells",#17
                     "Cardiomyocytes_Nppa+", #18
                     "Myeloid_cells", #19
                     "Endocardial_cells", #20
                     "Epicardial_cells", #21
                     "Myeloid_cells", #22
                     "Cardiac_fibroblasts", #23
                     "Lymphoid_cells", 
                     "Cardiac_fibroblasts"
                     )

names(new.cluster.ids) <- levels(harmony_object)
harmony_object <- RenameIdents(harmony_object, new.cluster.ids)
harmony_object@meta.data$anno_cell_type_lvl2 <- harmony_object@active.ident

## Reset cluster to seurat ID
Idents(object = harmony_object) <- "seurat_clusters"
```

```{r}
DimPlot(harmony_object, reduction = "umap", label = TRUE, repel = TRUE,raster=FALSE, group.by = "anno_cell_type_lvl2")
```


### Refine cell type cluster based on spatial annotation and marker expression



```{r}
## Endocardial cells
## First, let's reclassify endocardial cells based on the Qupath annotation we did above
harmony_object@meta.data <- harmony_object@meta.data %>%
  mutate("anno_cell_type_lvl2" = if_else(spatial_endoc == TRUE & anno_cell_type_lvl2 == "Endocardial_cells","Endocardial_cells",
                                         if_else(spatial_endoc == TRUE & anno_cell_type_lvl2 != "Endocardial_cells","Endocardial_cells",
                                          if_else(spatial_endoc == FALSE & anno_cell_type_lvl2 == "Endocardial_cells","Endothelial_cells",anno_cell_type_lvl2))))


# ## Add cells located within endocardium to annotations
# harmony_object@meta.data <- harmony_object@meta.data %>%
#   mutate("anno_cell_type_lvl2" = if_else(spatial_endoc == TRUE | anno_cell_type_lvl2 == "Endocardial_cells","Endocardial_cells",anno_cell_type_lvl2),
#          "anno_cell_type_lvl3" = if_else(spatial_endoc == TRUE | anno_cell_type_lvl3 == "Endocardial_cells_Npr3","Endocardial_cells_Npr3",anno_cell_type_lvl3))


# ## Separate out Lyve1 expressing cells belonging to lymphatic vessels
# lymph_cells <- subset(harmony_object, anno_cell_type_lvl2 == "Endocardial_cells" & Lyve1 > 0)
# 
# harmony_object@meta.data <- harmony_object@meta.data %>%
#   mutate("anno_cell_type_lvl2" = if_else(rownames(harmony_object@meta.data) %in% rownames(lymph_cells@meta.data),"Lymphatic_endothelial_cells",anno_cell_type_lvl2),
#          "anno_cell_type_lvl3" = if_else(rownames(harmony_object@meta.data) %in% rownames(lymph_cells@meta.data),"Lymphatic_endothelial_cells_Lyve1",anno_cell_type_lvl2))
```

```{r}
new_full_meta <- harmony_object@meta.data
## Plot cells that are 
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey") +
  geom_point(data = subset(new_full_meta,anno_cell_type_lvl2 == "Endocardial_cells"),color = "red") +
  geom_point(data = subset(new_full_meta,grepl("Myeloid_cells",anno_cell_type_lvl2)),color = "blue") +
  facet_wrap(~ sample_ID)
```


# Save object as seurat RDS

```{r}
saveRDS(harmony_object,
        file = "./output/molkart/molkart.harmony_seurat_object.mesmer_seg.rds")
```

```{r}
## If the object has already been computed
harmony_object <- readRDS(file
                          = "./output/molkart/molkart.harmony_seurat_object.mesmer_seg.rds")
```

# Generate Misty input

```{r}
misty_in <- harmony_object@meta.data %>%
  dplyr::select(sample_ID,timepoint,replicate,X_centroid,Y_centroid,Area,anno_cell_type_lvl2,nCount_RNA) 
misty_in$cell_ID <- rownames(misty_in)
```

```{r}
expression_plot_list <- list()

# final_samples <- c("sample_control_r1_s1","sample_control_r2_s1",
#                    "sample_4h_r1_s1","sample_4h_r2_s2",
#                    "sample_2d_r1_s1","sample_2d_r2_s1",
#                    "sample_4d_r1_s1","sample_4d_r2_s1")

samples <- c("sample_control_r1_s1","sample_4h_r1_s1",
             "sample_2d_r1_s1","sample_4d_r1_s1",
             "sample_control_r2_s1","sample_4h_r2_s2",
             "sample_2d_r2_s1","sample_4d_r2_s1")

for(this_sample in samples){
  pt_size <- 0.2
  cluster_of_int <- c(20,4)
  sample_object <- subset(misty_in,sample_ID == this_sample)
  
  highlight_plot <- ggplot(sample_object,aes(X_centroid,Y_centroid)) +
    geom_point(aes(color = anno_cell_type_lvl2),size = pt_size) +
    theme_classic() +
    labs(x = "Spatial 1",
         y = "Spatial 2",
         title = this_sample) +
    theme(axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "right") +
    scale_color_brewer(palette = "Set3")
  
      expression_plot_list[[this_sample]] <- highlight_plot
  
}

wrap_plots(expression_plot_list, nrow = 2, ncol = 4)  + plot_layout(guides = 'collect')
```

```{r}
misty_in_lowres <- misty_in %>%
  mutate("misty_cts" = anno_cell_type_lvl2)

write.table(misty_in_lowres,
            file = "./output/molkart/molkart.misty_celltype_table.mesmer_seg.lowres.tsv",
            sep="\t",
            col.names = TRUE,
            row.names = FALSE,
            quote = FALSE)
```


# Plot cell-type on spatial plot

```{r}
## Count percentage of each cell type in each sample
ct_prop <- harmony_object@meta.data %>%
  group_by(anno_cell_type_lvl2,sample_ID) %>%
  tally() %>%
  ungroup() %>%
  group_by(sample_ID) %>%
  mutate("prop" = n / sum(n))
```


```{r}
expression_plot_list <- list()

samples <- c("sample_control_r1_s1","sample_4h_r1_s1",
             "sample_2d_r1_s1","sample_4d_r1_s1",
             "sample_control_r2_s1","sample_4h_r2_s2",
             "sample_2d_r2_s1","sample_4d_r2_s1")

cluster_of_int <- 20
for(cluster_of_int in unique(harmony_object$anno_cell_type_lvl2)){
  print(cluster_of_int)
  for(this_sample in samples){
    pt_size <- 0.6
    sample_object <- subset(harmony_object,sample_ID == this_sample)
    meta <- sample_object@meta.data
    
    highlight_plot <- ggplot(meta,aes(Y_centroid,X_centroid)) +
      geom_point(data = subset(meta,!seurat_clusters %in% cluster_of_int),color = "darkgrey", size = pt_size) +
      # geom_point(data = subset(meta,seurat_clusters %in% cluster_of_int),aes(color = seurat_clusters), size = pt_size) +
      geom_point(data = subset(meta,anno_cell_type_lvl2 == cluster_of_int),color = "purple", size = 1) +
      theme_classic() +
      labs(x = "Spatial 1",
           y = "Spatial 2") +
      theme(axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "right")
    
    expression_plot_list[[this_sample]] <- highlight_plot
    }
  
  time_plot <- wrap_plots(expression_plot_list, nrow = 2, ncol = 4)  + plot_layout(guides = 'collect') + plot_annotation(cluster_of_int,theme=theme(plot.title=element_text(hjust=0.5)))
  
  print(time_plot)
}
```


# Save data formatted for input to spatial anndata object

```{r}
metadata <- harmony_object@meta.data %>%
  mutate("liana_cts" = anno_cell_type_lvl2)

coordinates <- metadata %>% select(X_centroid,Y_centroid)
counts <- t(as.matrix(harmony_object@assays$SCT@counts))

## Save coordinates
write.table(coordinates,
            "./output/molkart/mesmer_seg/harmony_object.coordinates_anndata.tsv",
            sep="\t",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)

## Save counts
write.table(counts,
            "./output/molkart/mesmer_seg/harmony_object.counts_anndata.tsv",
            sep="\t",
            col.names = TRUE,
            row.names = FALSE,
            quote = FALSE)

## Annotation
write.table(metadata,
            "./output/molkart/mesmer_seg/harmony_object.metadata_anndata.tsv",
            sep="\t",
            col.names = TRUE,
            row.names = FALSE,
            quote = FALSE)
```

```{r}
plot_density(harmony_object,features = c("Cd36","Vegfa","Wif1","Nppb"))
```

