---
title: "SeqIF - Quantification of myeloid cells in different cardiac regions"
author: "FloWuenne"
date: "2023-08-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(data.table)
library(ggbeeswarm)
library(here)

source(here("./code/functions.R"))
```


# Read data

We will read the cell type clustering output from pixie, which contains our cell-type labels and coordinates, as well as a table containing annotations in which region these cells fall, based on overlaying Cellpose mask centroids with Qupath annotations.

## Region annotation

```{r}
data_path <- "./data/seqIF_regions_annotations/"
cell_locations <-
    list.files(path = paste(data_path,"cell_locations",sep=""),pattern = "*.csv") %>%
    setNames(., .) %>%
    map_df(~fread(paste(paste(data_path,"cell_locations",sep=""),.,sep="/"), select= c("CellID","ROI")),
           .id = "sample") %>%
  mutate("fov" = gsub("annotated_|.csv","",sample))
  
region_measurements_full <- 
    list.files(path = paste(data_path,"ROI_measurements",sep=""),pattern = "*.txt") %>% 
    setNames(., .) %>%
    map_df(~fread(paste(paste(data_path,"ROI_measurements",sep=""),.,sep="/")),
           .id = "sample") %>%
  mutate("fov" = gsub(".txt","",sample)) 
colnames(region_measurements_full) <- gsub(" ","_",colnames(region_measurements_full))
colnames(region_measurements_full) <- gsub("\\^","",colnames(region_measurements_full))
colnames(region_measurements_full) <- gsub("\\µ","u",colnames(region_measurements_full))

## Sum all individual annotations per sample and region
region_measurements <- region_measurements_full %>%
  mutate("Area_um2" = if_else(fov %in% c("24h_86","4h_97"),Area_um2 * 0.235^2,Area_um2)) %>% ## Images that were post-registered using palom need to be size corrected, as their final pixel resolution is different
  group_by(fov,Name) %>%
  summarise("total_area_um2" = sum(Area_um2)) %>%
  ungroup()
```

## Pixie output

```{r}
pixie_cell_out <- fread("../data/SeqIF/pixie_cell_masks_0.05/cell_table_size_normalized_cell_labels.csv")
pixie_cell_out <- pixie_cell_out %>%
  separate(fov,into = c("time","sample"), remove = FALSE) %>%
  subset(cell_meta_cluster != "background") %>%
  mutate("CellID" = label)
```

## Merge data

```{r}
annotated_cells <- left_join(pixie_cell_out,cell_locations, by = c("fov","CellID"))
annotated_cells <- annotated_cells %>%
  subset(ROI != "Background")

## Set factor level for time
annotated_cells$time <- factor(annotated_cells$time,
                               levels = c("Control","4h","24h","48h"))
```

# Analyze data

## General statistics

```{r}
## What's the total number of cells we have classified with pixie?
print(paste("Total number of cells in the dataset:",nrow(annotated_cells),sep=" "))
```

## Cell-types per sample

```{r}
## Let's plot the number of cells per cell-type per sample
cells_per_sample <- annotated_cells %>%
  group_by(fov,time,cell_meta_cluster) %>%
  tally() %>%
  ungroup() %>%
  group_by(time,cell_meta_cluster) %>%
  summarise("n" = sum(n)) %>%
  ungroup() %>%
  group_by(time) %>%
  mutate("frac" = n / sum(n))
  

ggplot(cells_per_sample,aes(time,frac,fill = cell_meta_cluster)) +
  geom_bar(stat="identity", color = "black") +
  scale_fill_brewer(palette = "Set1")

# ggplot(cells_per_sample,aes(fov,n,fill = cell_meta_cluster)) +
#   geom_bar(stat="identity", color = "black") +
#   coord_flip() +
#   scale_fill_brewer(palette = "Set1")
```

## All immune cell types per region

```{r}
## Let's plot the number of cells per cell-type per sample
cells_per_region <- annotated_cells %>%
  subset(!ROI %in% c("Unclassified","Ignore")) %>%
  subset(!ROI %in% c("PathAnnotationObject","papillary_muscle","Lumen","Background",
                     "Ignore","RV_ignore","myocardium_control","remote_endocardium")) %>%
  group_by(ROI,fov,time,cell_meta_cluster) %>%
  tally()

cells_per_region_sub <- cells_per_region %>%
subset(grepl("Macro|Mono|Leuko|Neutro",cell_meta_cluster))

## Normalize cell numbers for area
region_measurements$ROI <- region_measurements$Name
cells_per_region_norm <- left_join(cells_per_region_sub,region_measurements, by = c("fov","ROI"))
cells_per_region_norm <- cells_per_region_norm %>%
  mutate("cells_per_mm2" = n / total_area_um2 * 1000000)  ## Normalize to square mm

ggplot(region_measurements,aes(fov,total_area_um2)) +
  geom_bar(stat="identity")

ggplot(cells_per_region_norm,aes(time,cells_per_mm2,fill = time)) +
  #geom_bar(stat="identity",position = "dodge", color = "black") +
  #geom_point(size = 3) +
  geom_boxplot() +
  theme_bw() +
  facet_grid(ROI ~ cell_meta_cluster,
             scales = "free_y") +
  scale_fill_brewer(palette = "Dark2")
```

## Ccr2+ mono/macros per region

```{r}
ccr2_monomacro <- cells_per_region_norm %>%
  subset(cell_meta_cluster == "Mono / Macros Ccr2+" | cell_meta_cluster == "Neutrophils") %>%
  subset(time %in% c("4h","24h","48h"))

ccr2_monomacro$ROI <- gsub("border_zone","Border zone",ccr2_monomacro$ROI)
ccr2_monomacro$ROI <- gsub("infarct_core","Infarct core",ccr2_monomacro$ROI)
ccr2_monomacro$ROI <- gsub("Epicardium","Epicard",ccr2_monomacro$ROI)
ccr2_monomacro$ROI <- gsub("Endocardium","Endocard",ccr2_monomacro$ROI)

ccr2_monomacro$ROI <- factor(ccr2_monomacro$ROI,
                             levels = c("Border zone","Infarct core","Epicard","Endocard"))

ccr2_mono_regions <- ggplot(ccr2_monomacro,aes(x = time,y = cells_per_mm2, fill = time)) +
    geom_beeswarm(size = 2, pch = 21, color = "black", aes(fill = time)) +
    stat_summary(
      fun.y = mean,
      geom = "errorbar",
      aes(ymax = ..y.., ymin = ..y..),
      width = 0.6,
      size = 1,
      color = "black") +
    labs(x = "Time",
         y = expression("Cells /"~mm^2)) + 
  #expression(paste("Mo / M",phi," per "~mm^2))
  facet_grid(cell_meta_cluster ~ ROI) +
  scale_fill_manual(values = time_palette[2:4]) + 
  theme_bw() +
  theme(axis.title = element_text(face="bold"),
        legend.position = "none") 

  
  # geom_signif(comparisons = list(c("MI_IZ","MI_remote")),
  #             tip_length = 0, annotation = "0.0057", y_position = 15.5) +
  # geom_signif(comparisons = list(c("MI_IZ","Control")),
  #             tip_length = 0, annotation = "0.0022", y_position = 16.5) +
  #  expand_limits(y = c(13, 17)) +
  # theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  # labs(x = "")
  
ccr2_mono_regions

save_plot(ccr2_mono_regions,
          file = "./plots/Figure3.ccr2_monomacro_regions.pdf",
          base_height = 3.5)
```



```{r}
subdata <- annotated_cells %>%
  subset(fov == "24h_86")
```

```{r}
ggplot(subdata,aes(X_centroid,Y_centroid)) +
  geom_point(data = subset(subdata,cell_meta_cluster != "Mono / Macros Ccr2+"),color = "darkgrey") +
  geom_point(data = subset(subdata,cell_meta_cluster == "Cardiomyocytes Ankrd1+"),color = "#EFCC00") +
  geom_point(data = subset(subdata,cell_meta_cluster == "Mono / Macros Ccr2+"),color = "purple")
  #geom_hex(data = subset(subdata,cell_meta_cluster == "Mono / Macros Ccr2+"),bins = 100) +
  scale_fill_viridis_c()

```

