---
title: "proteomics.impute_proteins"
author: "FloWuenne"
date: "2023-06-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(missForest)
library(here)
library(ggridges)
library(patchwork)
library(pheatmap)

source(here("./code/params_R.R"))
```

# Introduction

Here we will use the filtered protein list to impute missing proteins using iterated random forests via the [missForest package](https://cran.r-project.org/web/packages/missForest/index.html) as described by [Jin et al 2021](https://www.nature.com/articles/s41598-021-81279-4)
.

# Load data

First, we load the filtered data table produced in [Filter proteomics data](./proteomics.filter_proteomic_data.Rmd).

```{r}
prot <- fread("./output/proteomics.filtered_proteins.tsv")
genes_prot <- prot %>%
  select(Protein_Ids,Genes)

prot_form <- prot %>%
  select(-c(Protein_Group,Protein_Ids,Protein_Names,Genes))

rownames(prot_form) <- prot$Protein_Ids
```

# Impute proteins

No we will use the MissForest package to impute missing value in our protein matrix. We use the doParallel package to parallelize the imputation to speed up the process.

```{r}
## Logged before imputation
prot_form_log <- log2(prot_form)
doParallel::registerDoParallel(cores = 6) # set based on number of CPU cores
doRNG::registerDoRNG(seed = 123)
prot_form_RF_log <- missForest(prot_form_log, verbose = TRUE,
                           parallelize = 'forests')
```

```{r}
## Perform imputation using MissForest
prot_imp_log <- prot_form_RF_log$ximp
prot_imp_log$Protein_Ids <- rownames(prot_imp_log)
prot_imp_log <- left_join(prot_imp_log,genes_prot, by = "Protein_Ids")

## Add number of imputed values to table
imputed_val_rows <- rowSums(is.na(prot))
imputed_val_df <- data.frame("Genes" = prot$Genes,
                              "Protein_Ids" = prot$Protein_Ids,
                              "imputed_values" = imputed_val_rows)

prot_imp_log <- left_join(prot_imp_log,imputed_val_df, by =c("Genes","Protein_Ids"))
```

```{r}
## Transform data to long format
prot_imp_log_long <- prot_imp_log %>%
    pivot_longer(control_r1:MI_remote_r4,
               names_to = "sample",
               values_to = "exp") %>%
  mutate("group" = if_else(grepl("control",sample),"control",
                           if_else(grepl("MI_IZ",sample),"MI_IZ","MI_remote")))

## For each protein, add whether the value is imputed or observed
prot_long <- prot %>%
    pivot_longer(control_r1:MI_remote_r4,
               names_to = "sample",
               values_to = "exp") %>%
  mutate("imputed" = if_else(is.na(exp),"yes","no")) %>%
  select(Protein_Ids,Genes,sample,imputed)
  
prot_imp_log_long <- left_join(prot_imp_log_long,prot_long,by=c("Protein_Ids","Genes","sample"))

## Set order of groups
prot_imp_log_long$group <- factor(prot_imp_log_long$group,
                                levels = c("control","MI_remote","MI_IZ"))
```

Let's visualize the overall distribution of protein expression values after imputation per sample.

```{r}
## Plot distribution of all values
ggplot(prot_imp_log_long,aes(exp,sample,fill = group)) +
  geom_density_ridges2() +
  scale_fill_npg()
```

As we can see, protein expression distributions are very similar across samples and conditions.

Let's plot a sample correlation heatma, to see how samples correlate. We expect controls to correlate strongly with each other and more with remote endocardium then with MI endocardium.

```{r}
## Plot a correlation matrix of samples
prot_imp_log_matrix <- prot_imp_log %>%
  select(-c(Protein_Ids,Genes,imputed_values))

pheatmap(cor(prot_imp_log_matrix),
         border_color = "black",
         color = hcl.colors(50, "Inferno"),
         display_numbers = TRUE,
         cluster_cols = TRUE,
         cluster_rows = TRUE)
```

# Save imputed data

Let's save the imputed protein table for downstream analysis.

```{r}
write.table(prot_imp_log,
            file = "./output/proteomics.filt_imputed_proteins.tsv",
            sep = "\t",
            col.names = TRUE,
            row.names = FALSE,
            quote = FALSE)
```


# Compare to ND imputation

We want to also compare our iterative random forest imputation (RF) with an imputation from the [Perseus package that uses random draws from a normal distribution](http://www.coxdocs.org/doku.php?id=perseus:user:activities:matrixprocessing:imputation:replacemissingfromgaussian) (ND imputation).

First, let's load the Perseus imputed values:

```{r}
## Compare to Perseus ND imputation
perseus <- fread("./data/20230614_perseus_imputed.txt")

perseus_merge <- left_join(prot_imp_log,perseus,by=c("Protein_Ids","Genes"), suffix = c("_RF","_perseus"))
```

We will then reformat the data to be merged with our iterative random forest imputation to be able to compare them.


```{r}
perseus_merge_data <- perseus_merge %>%
  select(-c(Protein_Ids,Protein_Group,Protein_Names,Genes))
perseus_merge_data$Genes <- perseus_merge$Genes
perseus_merge_data$Protein_Ids <- perseus_merge$Protein_Ids

perseus_merge_data <- perseus_merge_data %>% 
  pivot_longer(control_r1_RF:MI_remote_r4_perseus,
               names_to = "sample",
               values_to = "exp") 
  
perseus_merge_data_mi <- perseus_merge_data %>%
  subset(grepl("MI_",sample)) %>%
  separate(sample,into = c("group","region","replicate","method"), sep = "_") %>%
  mutate("sample" = paste(group,region,replicate,sep="_")) %>%
  select(sample,method,exp,Genes,Protein_Ids)

perseus_merge_data_control <-perseus_merge_data %>%
  subset(grepl("control",sample)) %>%
  separate(sample,into = c("group","replicate","method"), sep = "_") %>%
  mutate("sample" = paste(group,replicate,sep="_")) %>%
  select(sample,method,exp,Genes,Protein_Ids)

perseus_merge_data_long <- rbind(perseus_merge_data_control,perseus_merge_data_mi)

perseus_merge_data_long <- left_join(perseus_merge_data_long,prot_long,by=c("Protein_Ids","Genes","sample"))

perseus_merge_data_wide <- perseus_merge_data_long %>%
  pivot_wider(names_from = "method", values_from = "exp") %>%
  mutate("diff" = RF - perseus)
```

## Estimate average difference per protein

We can then calculate the differences in imputed values between RF and ND imputation.

```{r}
avg_imp_diff <- perseus_merge_data_wide %>%
  group_by(Genes,Protein_Ids) %>%
  subset(imputed == "yes") %>%
  summarise("count" = n(),
            "mean" = mean(diff),
            "sd" = sd(diff)) %>%
  arrange(desc(mean))

missing_rate_rows <- rowSums(is.na(prot))
missing_rate_df <- data.frame("Genes" = prot$Genes,
                              "Protein_Ids" = prot$Protein_Ids,
                              "missing_rate" = missing_rate_rows)

avg_imp_diff <- left_join(avg_imp_diff,missing_rate_df, by =c("Genes","Protein_Ids"))
avg_imp_diff$missing_rate <- avg_imp_diff$missing_rate 
```

## Plot differences between imputation strategies

Let's first visualize the correlation between non-imputed and imputed values from RF and ND.

```{r}
## Plot correlation between imputed values for all samples
ggplot(perseus_merge_data_wide,aes(RF,perseus,color = imputed)) +
  geom_point() +
  facet_wrap(~sample) +
  labs(x = "Imputed values Random Forest",
       y = "Imputed values Perseus (ND)")
```

As we can see, non-imputed values perfectly correlate, as these are not changed by imputation. For imputed values, ND seems to impute lower values on average then RF. This is also represented by the next plot, looking at the distribution of differences across both methods.

```{r}
## Plot distribution of differences between imputation methods
ggplot(perseus_merge_data_wide,aes(diff)) +
  geom_histogram() +
  facet_wrap(~sample) +
  labs(x = "RF - ND imputed values")
```

Finally, let's plot the correlation between the average imputed values and the average observed values per protein for both methods. A method that imputes values close to the actual observed values should show some linear correlation, while a random method should show no correlation. 

```{r}
## Plot avg imputed expression values against the avg of the non-imputed per gene
perseus_merge_data_wide_imp_sum <- perseus_merge_data_wide %>%
  group_by(Genes,Protein_Ids,imputed) %>%
  summarise("avg_RF" = mean(RF),
            "avg_perseus" = mean(perseus))

## RF
comp_imp_wide_RF <- perseus_merge_data_wide_imp_sum %>%
  select(-avg_perseus) %>%
  pivot_wider(names_from = "imputed",
              values_from = "avg_RF",
              names_prefix = "RF_imp_")

rf_corr <- ggplot(comp_imp_wide_RF,aes(RF_imp_yes,RF_imp_no)) + 
  geom_point() +
  geom_smooth() +
  labs(x = "Imputed value average",
       y = "Observed value average",
       title = "Iterative random forest imputation")

## ND
comp_imp_wide_RF <- perseus_merge_data_wide_imp_sum %>%
  select(-avg_RF) %>%
  pivot_wider(names_from = "imputed",
              values_from = "avg_perseus",
              names_prefix = "perseus_imp_")

nd_corr <- ggplot(comp_imp_wide_RF,aes(perseus_imp_yes,perseus_imp_no)) + 
  geom_point() +
  geom_smooth() +
  labs(x = "Imputed value average",
       y = "Observed value average",
       title = "Perseus ND imputation")

## Plot correlation between average imputed values and observed values for both imputation methods
rf_corr + nd_corr
```

As the plots show, there is a clear linear correlation between the average imputed value from RF and the average observed values per protein, with RF on average imputing lower average values then the observed values, making it likely more conservative. ND imputation on the other hand shows no correlation with observed values at all, highlighting the random nature of the approach and suggesting that imputation using this method might induce more noise into the data.
