---
title: "endocardial_cell_definition"
author: "FloWuenne"
date: "2024-12-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(Seurat)
library(SeuratData)
#library(SeuratDisk)
library(data.table)
library(harmony)
library(here)
library(tidyverse)
library(Matrix)
library(sf)
library(sfheaders)
library(viridis)

source("./code/functions.R")

## Load annotated Seurat object
seurat_object <- readRDS("./output/molkart/molkart.seurat_object.rds")
```

## Use manual tissue annotation to improve endocardial cell labeling

Endocardial cells are very small and thin and form a monolayer at the inner lining of the heart. We noticed that our RNA marker Npr3 does not always reliably label endocardial cells and therefore will use manual annotation of the endocardial tissue regions to improve annotation of endocardial cells.

```{r}
## Function to use GeoJson mask from QuPath with Seurat metadata to check cell overlap to manual annotation region
check_mask_overlap <- function(polygon_sf,metadata,img_height,anno_name){
  points_sf <- st_as_sf(metadata, coords = c("X_centroid", "Y_centroid"), crs = NA)
  polygon_sf <- st_set_crs(polygon_sf, NA)
  
  mirrored_geometries <- st_geometry(polygon_sf) %>% 
    st_coordinates() %>%
    as.data.frame() %>%
    mutate("Y" = img_height - Y) %>%
    select(X,Y,L2)
  mirrored_geometries <- as.data.frame(mirrored_geometries)
  
  polygon_list <- split(mirrored_geometries[,c("X", "Y")], mirrored_geometries$L2)
  closed_polys <- list()
  # Iterate over polygons in polygon_list and add the first row to the end of each polygon
  for (i in 1:length(polygon_list)) {
    closed_polys[[i]] <- as.matrix(rbind(polygon_list[[i]], polygon_list[[i]][1,]))
  }
  
  # Convert the list of polygons to sf objects
  sf_objects <- lapply(closed_polys, function(polygon) {
    st_polygon(list(polygon))
  })
  
  # Create a MULTIPOLYGON object
  multipolygon <- st_multipolygon(sf_objects)
  
  ## Check which points fall withing the annotation
  points_within_polygons <- st_within(points_sf, multipolygon, sparse = FALSE)
  metadata[anno_name] <- points_within_polygons
  return(metadata)
}
```

```{r}
## Check which cells fall in the spatial endocardial region for each sample
all_vals <- list()
for(sample_id in unique(seurat_object@meta.data$sample_ID)){
  print(sample_id)
  image_sizes <- fread("./references/molkart.image_dimensions.csv")
  img_height <- subset(image_sizes,sample_ID == sample_id)$height
  
  polygon_sf <- st_read(paste("./annotations/molkart/heart_regions_r/",sample_id,".stack.geojson",sep=""))
  sample_meta <- subset(seurat_object@meta.data,sample_ID == sample_id)
  sample_vals <- list(polygon_sf,sample_meta,img_height)
  all_vals <- append(all_vals,list(sample_vals))
  #
}
#test <- check_mask_overlap(polygon_sf,sample_meta,img_height)
meta_list <- Map(function(x) check_mask_overlap(x[[1]], x[[2]], x[[3]],"in_endo"), all_vals)

# Combine all data frames into one
new_full_meta <- do.call(rbind, meta_list)
```

```{r}
## Which seurat cluster corresponds to endocardial cells (Npr3 positive)?
endocardial_cluster <- 15

## How many cells are in the endocardial cluster?
n_endo <- nrow(subset(new_full_meta,seurat_clusters == endocardial_cluster))

## How many of the transcriptionally defined endocardial cell cluster are spatially within the endocardium?
n_endo_spatial <- nrow(subset(new_full_meta,seurat_clusters == endocardial_cluster & in_endo == TRUE))

## How many segmented cells are total in the annotated regions?
anno_region_cells <- nrow(subset(new_full_meta, in_endo == TRUE))

## Let's classify cells that are Pecam1 positive within the endocardial region as endocardial (even if they are not Npr3 positive)
## add Pecam1 expression 
pecam1_exp <- seurat_object@assays$SCT@data["Pecam1",]
new_full_meta$Pecam1 <- pecam1_exp
npr3_exp <- seurat_object@assays$SCT@data["Npr3",]
new_full_meta$Npr3 <- npr3_exp

## How many cells are Pecam1 positive within the endocardial region?
pecam1_pos <- subset(new_full_meta,Pecam1 > 0 & in_endo == TRUE)

## Add another annotation column to metadata to classify all cells with Pecam1 > 0 within the endocardial region as endocardial cells
new_full_meta <- new_full_meta %>%
  mutate("spatial_endoc" = if_else((in_endo == TRUE & Pecam1 > 2) | (in_endo == TRUE & seurat_clusters == endocardial_cluster), TRUE,FALSE))

seurat_object@meta.data <- new_full_meta
```

### Plot results of spatial annotation for endocardial cells 

```{r}
## Plot cells that are transcriptionally labeled as endocardial cells
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey") +
  geom_point(data = subset(new_full_meta,seurat_clusters == endocardial_cluster),color = "red") +
  facet_wrap(~ sample_ID)

## Plot cells that are in cluster 15 but don't express Npr3
ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey") +
  geom_point(data = subset(new_full_meta,seurat_clusters == endocardial_cluster & Npr3 == 0),color = "red") +
  geom_point(data = subset(new_full_meta,seurat_clusters == endocardial_cluster & Npr3 >  0),color = "blue") +
  facet_wrap(~ sample_ID)

## Plot 
## 1: blue - originally transcriptionally labeled endocardial cells but not in endocardial region. 
## 2: red - transcriptionally labeled endocardial cells that were in endocardial region
## 3: purple - Cells that were not clustered with Npr3 positive cells but spatially fall within endocardium and are Pecam1 or Npr3 positive.
endocardial_spatial_annotation <- ggplot(new_full_meta,aes(X_centroid,Y_centroid)) +
  geom_point(color = "darkgrey", size = 0.5) +
  geom_point(data = subset(new_full_meta,spatial_endoc == TRUE & seurat_clusters == endocardial_cluster),color = "yellow", size = 0.5) +
  geom_point(data = subset(new_full_meta,spatial_endoc == TRUE & seurat_clusters != endocardial_cluster),color = "purple", size = 0.5) +
  geom_point(data = subset(new_full_meta,spatial_endoc == FALSE & seurat_clusters == endocardial_cluster),color = "blue", size = 0.5) +
  facet_wrap(~ sample_ID)

save_plot(endocardial_spatial_annotation,
          file = "./plots/molkart.endocardial_spatial_annotation.png",
          base_height = 6)
```

```{r}
new_seurat_object <- seurat_object
new_seurat_object@meta.data <- new_seurat_object@meta.data %>%
  mutate("anno_cell_type_lvl2" = if_else(spatial_endoc == TRUE & anno_cell_type_lvl2 == "Endocardial_cells","Endocardial_cells",
                                         if_else(spatial_endoc == TRUE & anno_cell_type_lvl2 != "Endocardial_cells","Endocardial_cells",
                                          if_else(spatial_endoc == FALSE & anno_cell_type_lvl2 == "Endocardial_cells","Endothelial_cells",anno_cell_type_lvl2))),
         "anno_cell_type_lvl3" = if_else(spatial_endoc == TRUE & anno_cell_type_lvl3 == "Endocardial_cells_Npr3","Endocardial_cells_Npr3",
                                         if_else(spatial_endoc == TRUE & anno_cell_type_lvl3 != "Endocardial_cells_Npr3","Endocardial_cells_Npr3",
                                         if_else(spatial_endoc == FALSE & anno_cell_type_lvl3 == "Endocardial_cells_Npr3","Lymphatic_endothelial_cells_Lyve1",anno_cell_type_lvl3))))

## Plot a dotplot of Npr3 across all cell types
library(viridis)
endocardial_cells_dotplot <- DotPlot(new_seurat_object, 
        features = c("Npr3"),
        group.by = "anno_cell_type_lvl2") + 
  RotatedAxis() +
  scale_colour_viridis(option="magma") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="black"))) + 
  labs(tags = "c") + theme(plot.tag = element_text(size = 20, face = "bold"))
endocardial_cells_dotplot

save_plot(endocardial_cells_dotplot,
          file = "./plots/molkart.endocardial_dotplot.png",
          base_height = 4)
```

