---
title: "figures.supplementary_figure_X.proteomics_qc"
author: "FloWuenne"
date: "2023-06-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

(future.globals.maxSize = 4000 * 1024^5)

library(tidyverse)
library(data.table)
library(here)
library(ggrepel)
library(patchwork)
library(ggVennDiagram)
library(viridis)
library(heatmaply)
library(ggsignif)
library(ggrepel)
library(ggbeeswarm)

source("./code/functions.R")
```

This supplementary figure contains QC plots for the proteomic data analysis.

```{r}
## load data
metadata_counts <- fread("./output/proteomics/proteomics.protein_missing_stats.tsv")
```


# Subfigure A - Total proteins identified per group

```{r}
## Summarize observations across group
metadata_counts_stat <- data_summary(metadata_counts,
                                     varname = "wo_cont", groupnames = c("group"))

avg_proteins <- mean(metadata_counts$wo_cont)
avg_proteins <- sd(metadata_counts$wo_cont)

metadata_counts_stat$group <- factor(metadata_counts_stat$group,
                                     levels = c("MI_IZ","MI_remote","control"))

## Barplot to show the number of proteins detected per group (excluding contaminants)
n_prot_plot <- ggplot(metadata_counts_stat,aes(group,wo_cont)) +
  geom_bar(aes(fill = group),stat="identity", color="black", 
           position=position_dodge()) +
  geom_point(data = metadata_counts, aes(group,wo_cont),
             fill = "white",pch = 21, color= "black",size = 4) +
  geom_errorbar(aes(ymin=wo_cont, ymax=wo_cont+sd), width=.2,
                 position=position_dodge(.9)) + 
  coord_flip() +
  scale_fill_manual(values = proteome_palette,
                    labels = c("Control","MI_remote","MI_IZ")) +
  labs(x = "",
       y = "Proteins detected") +
  theme(legend.position = "none")

n_prot_plot
```


# Subfigure B - % missing proteins vs number of nuclei collected


```{r}
metadata_counts$group <- factor(metadata_counts$group,
                                levels = c("control","MI_remote","MI_IZ"))

missingness_plot <- ggplot(metadata_counts,aes(missingness,n_nuclei, label = sample,
                           fill = group)) +
  geom_point(pch = 21, size = 5) +
  scale_fill_manual(values = proteome_palette,
                    labels = c("Control","MI_remote","MI_IZ")) +
  geom_text_repel() +
  labs(x = "% missing values",
       y = "# nuclei in tissue collected",
       fill = "Group") +
  theme(legend.position = "top")
missingness_plot

```

# Subfigure C - % variance explained by PCs

```{r}
pca_res <- readRDS("./output/proteomics/proteomics.pca_res.rds")

sdev <- pca_res$sdev[1:5]
var_explained <- data.frame(PC = paste0("PC",1:5),
                            var_expl = (sdev)^2/sum((sdev)^2)
                            )

pca_var <- ggplot(var_explained,aes(PC,var_expl)) +
  geom_bar(stat = "identity", color = "black", fill = "grey30") +
  geom_text(aes(label = paste(100*round(var_expl,2),sep = "")), vjust=1.4, color = "white", size =6) +
  labs(x = "Principal components",
       y = "% Variance explained")

pca_var
```

# Subfigure D - Venn Diagram of limma DE results

```{r}
## Load limma results and subset for significantly differential proteins between conditions
limma_res <- fread("./output/proteomics/proteomics.limma.full_statistics.tsv")
limma_res_sig <- subset(limma_res,adj.P.Val <= 0.05)

limma_mi_ctr_genes <- subset(limma_res_sig,analysis == "MI_IZ_vs_control")$gene
limma_remote_ctr_genes <- subset(limma_res_sig,analysis == "MI_remote_vs_control")$gene
limma_mi_remote_genes <- subset(limma_res_sig,analysis == "MI_IZ_vs_MI_remote")$gene
limma_venn_vec <- unique(limma_res_sig$gene)
limma_venn_df <- data.frame("MI_control" = limma_venn_vec %in% limma_mi_ctr_genes,
                      "remote_control" = limma_venn_vec %in% limma_remote_ctr_genes,
                      "MI_IZ_MI_remote" = limma_venn_vec %in% limma_mi_remote_genes)
limma_venn_df$gene <- limma_venn_vec

limma_venn_list <- list("MI_vs_control" = limma_mi_ctr_genes, 
                        "MI_remote_vs_control" = limma_remote_ctr_genes,
                        "MI_IZ_vs_MI_remote" = limma_mi_remote_genes)

## Get genes that are uniquely DE in MI vs control
mi_control_uniq <- subset(limma_venn_df,MI_control == TRUE & remote_control == FALSE & MI_IZ_MI_remote == FALSE)
mi_control_remote_shared <- subset(limma_venn_df,MI_control == TRUE & remote_control == FALSE & MI_IZ_MI_remote == TRUE)
mi_remote_uniq <- subset(limma_venn_df,MI_control == FALSE & remote_control == FALSE & MI_IZ_MI_remote == TRUE)

## Plot Venn diagram using ggVennDiagram

venn <- Venn(limma_venn_list)
data <- process_data(venn)
venn_diagram <- ggplot() +
  # change mapping of color filling
  geom_sf(fill = "lightgrey", data = venn_region(data), show.legend = FALSE) +  
  # adjust edge size and color
  geom_sf(color="black", size = 3, data = venn_setedge(data), show.legend = FALSE) +  
  # show set label in bold
  geom_sf_text(aes(label = name), size = 4, fontface = "bold", data = venn_setlabel(data)) +  
  # add a alternative region name
 geom_sf_label(aes(label = count), data = venn_region(data)) + 
  theme_cowplot() + 
  scale_x_continuous(expand = expansion(mult = .3)) 

venn_diagram$layers[[1]]$mapping <- aes(fill = name)

venn_diagram <- venn_diagram +
  scale_fill_manual(values = c(MI_vs_control = "white",
                               MI_vs_control..MI_remote_vs_control = "white",
                               MI_remote_vs_control = "white",
                               MI_vs_control..MI_IZ_vs_MI_remote = "white",
                               MI_IZ_vs_MI_remote = "white",
                               MI_remote_vs_control..MI_IZ_vs_MI_remote = "white",
                               MI_vs_control..MI_remote_vs_control..MI_IZ_vs_MI_remote = "white")) +
  theme(legend.position = "none") +
  theme(plot.background = element_rect(fill = NA, colour = "white", size = 0),
        axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.line =element_blank())

venn_diagram
```

# Volcano plot of MI_IZ vs control

```{r}
iz_control <- subset(limma_res, analysis == "MI_IZ_vs_control")
iz_control <- iz_control %>%
  arrange(desc(logFC))

prot_2label <- ""
prot_2label <- subset(iz_control, adj.P.Val < 0.05)[1:10,]$gene
prot_2label <- c(prot_2label,tail(subset(iz_control, adj.P.Val < 0.05),1)$gene)

iz_control <- iz_control %>%
  mutate("label_protein" = if_else(gene %in% prot_2label, gene,""))

## Volcano plot of IZ_remote vs control
iz_control_volc <- plot_pretty_volcano(iz_control, 
                                      pt_size = 2,
                                      plot_title = "MI_remote vs control",
                                      sig_col = "adj.P.Val",
                                      sig_thresh = 0.05,
                                      col_pos_logFC = proteome_palette[['control']],
                                      col_neg_logFC = proteome_palette[['MI_IZ']]) +
  geom_label_repel(max.overlaps = 10)

iz_control_volc
```

# Volcano plot of MI_remote vs control

```{r}
remote_control <- subset(limma_res, analysis == "MI_remote_vs_control")
remote_control <- remote_control %>%
  arrange(desc(logFC))

prot_2label <- subset(remote_control, adj.P.Val < 0.05)[1:13,]$gene
prot_2label <- c(prot_2label,tail(subset(remote_control, adj.P.Val < 0.05),1)$gene)
remote_control <- remote_control %>%
  mutate("label_protein" = if_else(gene %in% prot_2label, gene,""))

## Volcano plot of IZ_remote vs control
remote_control_volc <- plot_pretty_volcano(remote_control, 
                                      pt_size = 2,
                                      plot_title = "MI_remote vs control",
                                      sig_col = "adj.P.Val",
                                      sig_thresh = 0.05,
                                      col_pos_logFC = proteome_palette[['MI_remote']],
                                      col_neg_logFC = proteome_palette[['control']]) +
  geom_label_repel()

remote_control_volc
```


# Plot candidate protein expressions

```{r}
vsn_matrix <- fread("./output/proteomics/proteomics.vsn_norm_proteins.tsv")
#colnames(vsn_matrix)[1:11] <- paste("s",1:11,colnames(vsn_matrix)[1:11],sep="_")
protein_sub <- vsn_matrix  %>%
    dplyr::select(1:11,gene) %>%
    pivot_longer(1:11,names_to = "sample", values_to = "exp") %>%
    mutate("group" = if_else(grepl("control",sample),"control",
                                   if_else(grepl("MI_IZ",sample),
                                           "MI_IZ","MI_remote"))
    )

protein_sub$group <- factor(protein_sub$group,
                              levels = c("control","MI_remote","MI_IZ"))

goi <- "Vwf"
plot_proteomics_boxplot(norm_table = protein_sub,
                        protein = goi) +
  geom_signif(comparisons = list(c("MI_IZ","MI_remote")),
              tip_length = 0, annotation = subset(limma_res,
                                                  analysis == "MI_IZ_vs_MI_remote" &
                                                    gene == goi)$adj.P.Val)
```

```{r}
## "MI_IZ_vs_control"     "MI_remote_vs_control" "MI_IZ_vs_MI_remote"
strongest_proteins <- limma_res_sig %>%
  group_by(analysis) %>%
  arrange(desc(logFC)) %>%
  subset(analysis == "MI_IZ_vs_MI_remote")

vsn_matrix_long <- vsn_matrix %>%
  pivot_longer(control_r1:MI_remote_r4,
               names_to = "sample",
               values_to = "exp")

vsn_matrix_t <- vsn_matrix %>%
  column_to_rownames(var="gene")
vsn_matrix_t <- as.data.frame(t(vsn_matrix_t))
vsn_matrix_t <- vsn_matrix_t %>%
  dplyr::select(strongest_proteins[1:40,]$gene)


# library(heatmaply)
# exp_heatmap <- heatmaply(vsn_matrix_t, plot_method = "ggplot")
# exp_heatmap
```


# Assemble final figure

```{r,fig.height = 8, fig.width = 10}
supp_figure_X <- (n_prot_plot |  missingness_plot) / (pca_var | venn_diagram)

supp_figure_X <- supp_figure_X + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 25)) +
  theme(plot.background = element_rect(fill = "white"))

supp_figure_X

save_plot(filename = "./figures/Supplementary_figure_X.proteomics.png",
          plot = supp_figure_X,
          base_height = 8)


save_plot(filename = "./figures/Supplementary_figure_X.proteomics.eps",
          plot = supp_figure_X,
          base_height = 8)
```


