---
title: "Supplementary Figure 2"
author: "FloWuenne"
date: "2023-06-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RColorBrewer)
library(patchwork)
library(vroom)
library(ggpubr)

source("./code/functions.R")
```

# Supplementary Figure 2 - Quality control (QC) of Molecular Cartography data

The data used in this plot was calculated using the following scripts:

- [molkart.QC_spots](./analysis/molkart.QC_spots.Rmd)
* [Quantification of RNA spots on tissue](./analysis/molecular_cartography_python/molkart.count_spots_on_tissue.ipynb)

## A) Correlation between technical replicates


```{r}
merge_tx_sums <- vroom("./output/molkart/tx_abundances_per_slide.tsv")

merge_tx_sums_split <- merge_tx_sums %>%
  separate(sample_ID, into = c("string","time","replicate"))

# Replace replicate by slide ID
merge_tx_sums_split$replicate <- gsub("r1","Biol. repl. 1",merge_tx_sums_split$replicate)
merge_tx_sums_split$replicate <- gsub("r2","Biol. repl. 2",merge_tx_sums_split$replicate)
merge_tx_sums_split$time <- gsub("2d","2 days",merge_tx_sums_split$time)
merge_tx_sums_split$time <- gsub("4d","4 days",merge_tx_sums_split$time)

# Set order of time 
merge_tx_sums_split$time <- factor(merge_tx_sums_split$time,
                                   levels = c("control","4h","2 days","4 days"))

tx_correlation_plot <- ggplot(merge_tx_sums_split,aes(log10(total_tx_rep1),log10(total_tx_rep2))) +
    geom_point(aes(color = time)) +
    geom_smooth(method = "lm", color = "black") +
    labs(x = "log10(spots) - Slide 1",
         y = "log10(spots) - Slide 2") +
  stat_cor(aes(label = ..r.label..), method = "spearman") +
  facet_grid(replicate ~ time) +
  scale_color_brewer(palette = "Dark2") + 
  theme(strip.text = element_text(face = "bold", color = "black", size = 14),
        strip.background = element_rect(fill = "lightgrey", linetype = "solid",
                                        color = "black", linewidth = 0.8),
        axis.title = element_text(face="bold"),
        legend.position = "none"
        ) +
  panel_border()


tx_correlation_plot
```


## B)  Transcripts per uM of tissue


```{r}
spots_tissue <- vroom("./output/molkart/molkart.spots_per_tissue.tsv")
spots_tissue <- spots_tissue %>%
  separate(sample, into =c("sample","time","replicate","slide"), sep = "_")

spots_tissue$time <- factor(spots_tissue$time,levels = c("control","4h","2d","4d"))
spots_tissue <- spots_tissue %>%
  mutate("spots_per_mm2" = spots_per_um2 * 100)

spots_per_um <- ggbarplot(spots_tissue, 
                          x = "time", 
                          y = "spots_per_mm2",
                          add = c("mean", "dotplot"), 
                          fill = "time", color = "black",
                          palette = "Dark2") +
   labs (x = "Time",
         y = expression(bold(paste("10 000 Spots / ",m,m^2, sep="")))) +
  font("xlab", size = 16, color = "black", face = "bold") +
  font("ylab", size = 16, color = "black", face = "bold") +
    scale_y_continuous(
    # don't expand y scale at the lower end
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_minimal_hgrid() +
  theme(axis.title = element_text(face="bold")) +
  rremove("legend")

 spots_per_um
 
 avg_spots_per_um = mean(spots_tissue$spots_per_mm2)
 avg_spots_per_um
```

## C) Principal component analysis between samples


```{r} 
pcs <- vroom("./output/molkart/pca_spots.tsv")
pcs$time <- factor(pcs$time,
                   levels = c("control","4h","2d","4d"))
```


```{r}
pca_plot <- ggplot(pcs,aes(PC1,PC2,label = label)) +
  geom_point(size = 5, aes(color = time, shape = slide)) +
  scale_color_brewer(palette = "Dark2",labels = c("control","4 hours","2 days","4 days")) +
  labs(color = "Time",
       shape = "Slide")
pca_plot
```

# Assemble final figure

```{r,fig.height = 8, fig.width = 10}
supp_figure_2 <- tx_correlation_plot / (spots_per_um | pca_plot)

supp_figure_2 <- supp_figure_2 + 
  plot_annotation(tag_levels = 'a') & 
  theme(plot.tag = element_text(size = 25)) +
  theme(plot.background = element_rect(fill = "white"))

supp_figure_2

save_plot(filename = "./plots/Supplementary_figure_2.png",
          plot = supp_figure_2,
          base_height = 8)


save_plot(filename = "./plots/Supplementary_figure_2.eps",
          plot = supp_figure_2,
          base_height = 8)
```
