---
title: "molkart.celltype_analysis"
author: "FloWuenne"
date: "2024-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r}
library(Seurat)
library(here)
library(patchwork)
library(Nebulosa)
library(here)
library(reactable)
library(ggdark)

source("../code/functions.R")
```

# Introduction

In this analsysis, we will perform more granular analysis of the cell types we identified in Molecular Cartography and describe what kind of cells are present in the tissue.

# Load data

First we load our processed Seurat object that contains all cells and cell-type annotations.

```{r}
seurat_object <- readRDS(file = "../output/molkart/molkart.seurat_object.rds")

immune_cells <- subset(seurat_object,
                       anno_cell_type_lvl2 %in% c("Myeloid_cells","Lymphoid_cells"))
```

Next, let's plot a UMAP plot for all three levels of cell-type annotation we did.

```{r}
celltypes_lvl3 <- DimPlot(seurat_object, reduction = "umap", label = TRUE, repel = TRUE,raster=FALSE, group.by = "anno_cell_type_lvl3") + NoLegend() +
  labs (title = "Level 3 cell type annotations")

celltypes_lvl3
```

# Subcluster immune cells

```{r}
# Identify variable features
immune_cells <-  SCTransform(immune_cells,verbose = FALSE)

# Rerun PCA and UMAP 
immune_cells <- RunPCA(immune_cells, approx=FALSE)

# Run UMAP
immune_cells <- RunUMAP(immune_cells, dims = 1:25)

# Subcluster the cells in immune_cells
immune_cells <- FindNeighbors(immune_cells, dims = 1:25, verbose = FALSE)
immune_cells <- FindClusters(immune_cells, verbose = FALSE, resolution = 0.2)

# Plot the UMAP with the new clusters
DimPlot(immune_cells, reduction = "umap", label = TRUE, repel = TRUE)

subcluster_umap <- DimPlot(immune_cells, reduction = "umap", label = TRUE, repel = TRUE, group.by = "anno_cell_type_lvl3")
```

```{r}
immune_cells <- PrepSCTFindMarkers(immune_cells)
finegrained_markers <- FindAllMarkers(immune_cells, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
top3_markers_immunecells <- finegrained_markers %>% 
  group_by(cluster) %>% 
  top_n(2, avg_log2FC)
```

```{r}
# Show marker table using an interactive table viewer like reactable
reactable::reactable(finegrained_markers,sortable = TRUE,
                     searchable = TRUE, defaultPageSize = 10)
```

Let's visualize expression of "marker genes" for different immune cell types using Nebulosa:

```{r}
## Plot defined markers for each cell type (Let's make two separate plots for nicer rendering in workflowR pages)
plot_density(immune_cells, c("Csf3r","Ccl2","C1qa","Arg1"))
```

```{r}
plot_density(immune_cells,c("Cd3e","Ighm","Cd74"))
```


Let's also plot a heatmap for markers for these more fine grained cell types:

```{r}
library(viridis)
immune_cells_dotplot <- DotPlot(immune_cells, 
        features = c("Csf3r","Ccl2","C1qa","Arg1","Cd3e","Ighm","Cd74"),
        group.by = "anno_cell_type_lvl3") + 
  RotatedAxis() +
  scale_colour_viridis(option="magma") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="black")))
immune_cells_dotplot
```

# Spatial plot of subclustering results for immune cells

```{r}
## Perform similar rotation correction on images as in main Supplementary Figure 5
metadata <- immune_cells@meta.data

## Set sample order for patchwork
sample_names <- c("sample_4h_r1_s1",
             "sample_2d_r1_s1",
             "sample_4h_r2_s2",
             "sample_2d_r2_s1")

new_metadata <- data.frame()
# For each sample in sample_names, apply a rotation function to x and y positions in metadata
for(sample in sample_names){
  meta_sub <- subset(metadata,sample_ID == sample)
  if(sample == "sample_control_r1_s1"){ # rotate 90 degrees counter clockwise
    y_orig <- meta_sub$Y_centroid
    x_orig <- meta_sub$X_centroid 
    meta_sub$X_centroid <- y_orig * -1
    meta_sub$Y_centroid <- x_orig
  }else if(sample == "sample_control_r2_s1"){ # rotate 90 degrees counter clockwise
    meta_sub$X_centroid <- meta_sub$X_centroid * -1
    meta_sub$Y_centroid <- meta_sub$Y_centroid *  1
  }else if(sample == "sample_4h_r1_s1"){ # rotate 90 degrees counter clockwise
    meta_sub$X_centroid <- meta_sub$X_centroid * -1
    meta_sub$Y_centroid <- meta_sub$Y_centroid *  1
  }else if(sample == "sample_4h_r2_s2"){ # rotate 90 degrees counter clockwise
    y_orig <- meta_sub$Y_centroid
    x_orig <- meta_sub$X_centroid 
    meta_sub$X_centroid <- y_orig * -1
    meta_sub$Y_centroid <- x_orig
  }else if(sample == "sample_2d_r1_s1"){ # rotate 90 degrees counter clockwise
    meta_sub$X_centroid <- meta_sub$X_centroid * -1
    meta_sub$Y_centroid <- meta_sub$Y_centroid * -1
  }else if(sample == "sample_4d_r2_s1"){ # rotate 90 degrees counter clockwise
    y_orig <- meta_sub$Y_centroid 
    meta_sub$Y_centroid <- meta_sub$X_centroid * -1
    meta_sub$X_centroid <- y_orig * 1
  }
  new_metadata <- rbind(new_metadata,meta_sub)
}
new_metadata
```


```{r}
## Make a spatial plot for each sample and organize them using patchwork
sample_plots <- lapply(sample_names, function(sample) {
  meta_sub <- subset(new_metadata,sample_ID == sample)
  meta_sub$anno_cell_type_lvl3 <- gsub("_"," ",meta_sub$anno_cell_type_lvl3)
  
  # Generate the plot
  ## Add a line of blank space on top of the plot
  p <- ggplot(meta_sub, aes(x = X_centroid, 
                            y = Y_centroid,
                            color = anno_cell_type_lvl3)) +
    dark_theme_void() +
    geom_point(size = 0.6) +
    theme(legend.position = "right",
          legend.title = element_text(size = 20),
          legend.text = element_text(size = 16)) + # Hide legend or adjust as needed
    guides(color = guide_legend(title = "Cell types",override.aes = list(size = 6))) +
    labs(title = "")
  
  ## Manually add tags and titles to plots for visual improvements
  if(sample == "sample_control_r1_s1"){
    p <- p + 
      labs(tag = "a",
           title = "Control",
           y = "Replicate 1") + 
      theme(plot.tag = element_text(size = 20, face = "bold"),
            plot.title = element_text(hjust=0.5, size = 20, face = "bold"),
            axis.title.y = element_text(size = 20,angle = 90, face = "bold")
            )
  } else if(sample == "sample_4h_r1_s1"){
    p <- p + 
      labs(title = "4 hours") + 
      theme(plot.tag = element_text(size = 20, face = "bold"),
            plot.title = element_text(hjust=0.5, size = 20, face = "bold"))
  } else if(sample == "sample_2d_r1_s1"){
    p <- p + 
      labs(title = "2 days") + 
      theme(plot.tag = element_text(size = 20, face = "bold"),
            plot.title = element_text(hjust=0.5, size = 20, face = "bold"))
  } else if(sample == "sample_4d_r1_s1"){
    p <- p + 
      labs(title = "4 days") + 
      theme(plot.tag = element_text(size = 20, face = "bold"),
            plot.title = element_text(hjust=0.5, size = 20, face = "bold"))
  } else if (sample == "sample_control_r2_s1"){
    p <- p +
      labs(y = "Replicate 2") + 
      theme(axis.title.y = element_text(size = 20,angle = 90, face = "bold")
            )
  }
  
  
  # Return the plot
  return(p)
})

spatial_plot_immune_cells <- wrap_plots(sample_plots,ncol = 2, nrows = 2) + plot_layout(guides = "collect") & theme(legend.position = 'right')
spatial_plot_immune_cells
```
# Temporal changes in immune cell subclusters (R3-8)

```{r}
cell_type_proportions_per_sample <- seurat_object@meta.data %>%
  group_by(anno_cell_type_lvl3,timepoint,sample_ID) %>%
  tally() %>%
  ungroup() %>%
  group_by(timepoint,sample_ID) %>%
  mutate("percent" = n / sum(n) * 100) %>%
  ungroup()

## Check that percentage sum up to 100 per sample
percent_check <- cell_type_proportions_per_sample %>%
  group_by(sample_ID) %>%
  summarise("sum" = sum(percent))

## Replace underscores in cell names with spaces
cell_type_proportions_per_sample$anno_cell_type_lvl3 <- gsub("_"," ",cell_type_proportions_per_sample$anno_cell_type_lvl3)

## Set order of cell types in plot from most abundant to least abundant
mean_percent <- cell_type_proportions_per_sample %>%
  group_by(timepoint,anno_cell_type_lvl3) %>%
  summarise("mean_percent" = mean(percent)) %>%
  ungroup() %>%
  group_by(timepoint) %>%
  arrange(desc(mean_percent))

cell_type_proportions_per_sample$anno_cell_type_lvl3 <- factor(cell_type_proportions_per_sample$anno_cell_type_lvl3,
       levels = subset(mean_percent,timepoint == "control")$anno_cell_type_lvl3)

## Calculate mean cell number per timepoint
mean_percent <- cell_type_proportions_per_sample %>%
  mutate("anno_cell_type_lvl3" = if_else(grepl("Myeloid cells",anno_cell_type_lvl3),"Myeloid_cells",anno_cell_type_lvl3)) %>% ## Merge myeloid cells into one group for calculations
  group_by(timepoint,anno_cell_type_lvl3) %>%
  summarise("mean_percent" = mean(percent)) %>%
  ungroup()

## Subset for only immune cells
cell_type_proportions_per_sample <- cell_type_proportions_per_sample %>%
  filter(anno_cell_type_lvl3 %in% c("DC Cd74","Lymphoid cells Cd3e","Mono/Macros Arg1+","Mono/Macros C1qa","Mono/Macros Ccl2",
                                    "Neutrophils Csf3r"))

ct_time_barplot <- ggplot(cell_type_proportions_per_sample,aes(timepoint,percent,fill = anno_cell_type_lvl3)) +
  geom_bar(position='dodge', stat='summary', fun='mean', color = "black") +
  # Add points for each fill group on top of bars
  geom_point(position=position_dodge(width=0.9), aes(group=anno_cell_type_lvl3), size=2.5, shape=21, fill="white", color="black") +
  #scale_fill_manual(values = named_colors, name = "Cell types") +
  labs(x = " Timepoint",
       y = "% total cells",
       fill = "Cell types") +
  theme_minimal() +
  theme(legend.text=element_text(size=18),
        legend.title = element_text(size=20),
        legend.position = "right",
        axis.title = element_text(size=18),
        axis.text  = element_text(size=18))


ct_time_barplot <- ct_time_barplot + theme(plot.tag = element_text(size = 20, face = "bold"))
ct_time_barplot
```
